rgbif: R client for working with GBIF species occurrence data

### 2016-01-07 manuscript.Rmd
manuscript updates, more text around GBIF uses in other related pkgs, map eg

* Get a single numeric count of occurrenes - `occ_count()`
* Search for occurrences - `occ_search()`
* Get occurrences by occurrence identifier - `occ_get()`
* Get occurrence metadata - `occ_metadata()`

@@

#### Cleaning data

@@

@@

@@

The `rgbif` package is relatively stable now, and should not have
many breaking changes unless necessitated due to changes in the
GBIF API.

@@

This project was supported in part by the Alfred P Sloan
Foundation (Grant 2013-6-22).

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
[github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
can be found at [github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
under the open and permissive MIT license.

@@

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html

>>>>

* Get a single numeric count of occurrenes - `occ_count()`
* Search for occurrences - `occ_search()`
* A simplified and optimized version of `occ_search()` - `occ_data()`
* Get occurrences by occurrence identifier - `occ_get()`
* Get occurrence metadata - `occ_metadata()`

@@

#### Data cleaning

@@

## Mapping

An obvious downstream use case for species occurrence data is to map the
data. `rgbif` per se is largely not concerned with making this easier,
although we do have a simple wrappar around `ggplot2` to make it easy to
get a quick plot of occurrence data. For example, here we plot 100
occurrences for _Puma concolor_.

```{r fig.width=8, fig.height=4}
key <- name_backbone(name='Puma concolor')$speciesKey
dat <- occ_search(taxonKey = key, limit = 100, hasCoordinate = TRUE)
gbifmap(dat$data)
```

Another package, [spoccutils][spoccutils]

## GBIF data in other R packages

We discuss usage of GBIF data in other R packages throughout the
manuscript, but provide a synopsis here for clarity.

### taxize

Some of the GBIF taxonomic services are also avialable in [taxize][taxize],
an R package that focuses on getting data from taxonomic data sources on
the web. For example, with `get_gbifid()` one can get GBIF IDs used for
a set of taxonomic names - then use those IDs in other functions in `taxize`
to get additional information, like taxonomically downstream children.

### spocc

GBIF occurrence data is available in the R package [spocc][spocc] via `rgbif`.
`spocc` is a unified interface for fetching species occurrence data from
many sources on the web. For example, a user can collect occurrence data
from GBIF, iDigBio, and iNaturalist, and easily combine them, then use other
pacakages to clean and vizualize the data.

@@

The following are three use cases for `rgbif`: niche modelling,
spatial change in biodiversity, and distribution mapping.

@@

The `rgbif` package is relatively stable, and should not have
many breaking changes unless necessitated due to changes in the
GBIF API.

@@

This project was supported in part by the Alfred P Sloan
Foundation (Grant 2013-6-22), and the Helmsley Foundation
(Grant XXXXXX).

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
[github.com/sckott/msrgbif](https://github.com/sckott/msrgbif). Software
can be found at [https://github.com/ropensci/rgbif](https://github.com/ropensci/rgbif),
under an MIT license.

@@

[mit]: http://choosealicense.com/licenses/mit/
[ghrgbif]: https://github.com/ropensci/rgbif
[cranrgbif]: https://cran.rstudio.com/web/packages/rgbif/
[gbifdocs]: http://www.gbif.org/developer/summary
[gbifissues]: http://gbif.github.io/gbif-api/apidocs/org/gbif/api/vocabulary/OccurrenceIssue.html
[taxize]: https://github.com/ropensci/taxize
[spocc]: https://github.com/ropensci/spocc
[spoccutils]: https://github.com/ropensci/spocc

### 2016-01-07 manuscript.tex
manuscript updates, more text around GBIF uses in other related pkgs, map eg

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2490).

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrenes - \texttt{occ\_count()}
\item
  Search for occurrences - \texttt{occ\_search()}
\item
  Get occurrences by occurrence identifier - \texttt{occ\_get()}
\item
  Get occurrence metadata - \texttt{occ\_metadata()}
\end{itemize}

@@

\paragraph{Cleaning data}\label{cleaning-data}

@@

@@

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-25-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-30-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-37-1.pdf}
\caption{}
\end{figure}

@@

The \texttt{rgbif} package is relatively stable now, and should not have
many breaking changes unless necessitated due to changes in the GBIF
API.

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
\href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
Software can be found at
\href{https://github.com/ropensci/rgbif}{github.com/ropensci/rgbif},
under the open and permissive MIT license.

>>>>

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2580).

@@

\begin{itemize}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Get a single numeric count of occurrenes - \texttt{occ\_count()}
\item
  Search for occurrences - \texttt{occ\_search()}
\item
  A simplified and optimized version of \texttt{occ\_search()} -
  \texttt{occ\_data()}
\item
  Get occurrences by occurrence identifier - \texttt{occ\_get()}
\item
  Get occurrence metadata - \texttt{occ\_metadata()}
\end{itemize}

@@

\paragraph{Data cleaning}\label{data-cleaning}

@@

\subsection{Mapping}\label{mapping}

An obvious downstream use case for species occurrence data is to map the
data. \texttt{rgbif} per se is largely not concerned with making this
easier, although we do have a simple wrappar around \texttt{ggplot2} to
make it easy to get a quick plot of occurrence data. For example, here
we plot 100 occurrences for \emph{Puma concolor}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{key <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Puma concolor'}\NormalTok{)$speciesKey}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{key, }\DataTypeTok{limit =} \DecValTok{100}\NormalTok{, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{)}
\KeywordTok{gbifmap}\NormalTok{(dat$data)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-20-1.pdf}
\caption{}
\end{figure}

Another package, \href{https://github.com/ropensci/spocc}{spoccutils}

\subsection{GBIF data in other R
packages}\label{gbif-data-in-other-r-packages}

We discuss usage of GBIF data in other R packages throughout the
manuscript, but provide a synopsis here for clarity.

\subsubsection{taxize}\label{taxize}

Some of the GBIF taxonomic services are also avialable in
\href{https://github.com/ropensci/taxize}{taxize}, an R package that
focuses on getting data from taxonomic data sources on the web. For
example, with \texttt{get\_gbifid()} one can get GBIF IDs used for a set
of taxonomic names - then use those IDs in other functions in
\texttt{taxize} to get additional information, like taxonomically
downstream children.

\subsubsection{spocc}\label{spocc}

GBIF occurrence data is available in the R package
\href{https://github.com/ropensci/spocc}{spocc} via \texttt{rgbif}.
\texttt{spocc} is a unified interface for fetching species occurrence
data from many sources on the web. For example, a user can collect
occurrence data from GBIF, iDigBio, and iNaturalist, and easily combine
them, then use other pacakages to clean and vizualize the data.

@@

The following are three use cases for \texttt{rgbif}: niche modelling,
spatial change in biodiversity, and distribution mapping.

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-26-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
\caption{}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-38-1.pdf}
\caption{}
\end{figure}

@@

The \texttt{rgbif} package is relatively stable, and should not have
many breaking changes unless necessitated due to changes in the GBIF
API.

@@

All scripts and data used in this paper can be found in the permanent
data archive Zenodo under the digital object identifier (DOI). This DOI
corresponds to a snapshot of the GitHub repository at
\href{https://github.com/sckott/msrgbif}{github.com/sckott/msrgbif}.
Software can be found at
\href{https://github.com/ropensci/rgbif}{\url{https://github.com/ropensci/rgbif}},
under an MIT license.
